\documentclass[10pt,conference]{IEEEtran}
\IEEEoverridecommandlockouts

% --- packages ---
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{positioning,fit,arrows.meta,calc}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18} % 標準設定（目盛り方向などは変更しない）
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{cite}
\usepackage{balance}
\usepackage[hidelinks]{hyperref} % ← できるだけ最後に

% --- title/author ---
\title{LPDDR+FeRAM for Mobile Edge AI:\\
Chiplet/SiP Integration as a Practical Path}

\author{%
  \IEEEauthorblockN{Shinichi Samizo}%
  \IEEEauthorblockA{Independent Semiconductor Researcher\\
  Former Engineer, Seiko Epson Corporation\\
  Email: \href{mailto:shin3t72@gmail.com}{shin3t72@gmail.com}\\
  GitHub: \href{https://github.com/Samizo-AITL}{Samizo-AITL}}%
}

\begin{document}
\maketitle

% ===== Abstract =====
\begin{abstract}
Low-power DRAM (LPDDR) is the dominant main memory for mobile edge AI accelerators, balancing bandwidth and energy efficiency. 
However, LPDDR remains volatile and incurs standby power due to periodic refresh. 
Ferroelectric RAM (FeRAM), based on HfO$_2$, provides non-volatility, low-voltage operation, and fast rewriting, making it suitable as an assistive memory for checkpointing and state retention. 
Because monolithic LPDDR+FeRAM co-fabrication is infeasible due to process--temperature mismatch, this work proposes \textbf{chiplet-level LPDDR+FeRAM integration} using SiP/PoP packaging. 

System-level analysis, based on representative LPDDR5/5X and HfO$_2$-FeRAM parameters from prior silicon reports, shows that FeRAM chiplets can reduce standby power by up to 20\%, shorten resume latency from $\sim$10~ms (baseline LPDDR) to sub-ms range ($<$500~$\mu$s), and improve overall energy efficiency by 15--25\% under mobile edge AI workloads such as on-device inference, federated learning, and AR/VR. 
These results are derived from analytical modeling rather than prototype measurements, positioning the study as a \emph{design framework exploration} rather than a device demonstration.

The proposed architecture is coordinated by the \textbf{SystemDK co-design framework}, which manages checkpoint and refresh-offload policies across architecture, package, and runtime layers. 
\textbf{Target implementation nodes:} SoC at 5--3\,nm (FinFET/GAAFET), LPDDR5/5X at 1$\alpha$--1$\gamma$ DRAM nodes ($\sim$14--10\,nm), and a 28--22\,nm CMOS FeRAM chiplet integrated via SiP/PoP. 
This approach highlights a near-term, manufacturable path toward energy-efficient and responsive memory subsystems, while providing an educational reference for heterogeneous memory integration.
\end{abstract}

% ===== Sections =====
\input{sections/introduction}
\input{sections/device_process}
\input{sections/results}
\input{sections/outlook}
\input{sections/conclusion}

% ===== References =====
\balance
\vspace{1\baselineskip} % タイトルの上だけ余白を追加
\bibliographystyle{IEEEtran}
\bibliography{references}

% ===== Author Biography =====
% References と Bio の間を少しだけ詰める（必要に応じて -0.5～-1.0ex 程度で調整）
\vspace{-0.5\baselineskip}
\begingroup\small
\begin{IEEEbiographynophoto}{Shinichi Samizo}
received the M.S. degree in Electrical and Electronic Engineering from Shinshu University, Japan.
He worked at Seiko Epson Corporation as an engineer in semiconductor memory and mixed-signal device development, and also contributed to inkjet MEMS actuators and PrecisionCore printhead technology.
He is currently an independent semiconductor researcher focusing on process/device education, memory architecture, and AI system integration.

\textbf{Contact:} \href{mailto:shin3t72@gmail.com}{shin3t72@gmail.com},
\href{https://github.com/Samizo-AITL}{Samizo-AITL}
\end{IEEEbiographynophoto}
\endgroup

\end{document}
