\begin{abstract}
\textbf{和文要旨}:~
本稿は，Pbフリー・低温プロセス整合の\emph{静電薄膜MEMSアクチュエータ}を用いた
バイオインクジェットヘッドの設計・評価を通じて，
「壊さない精密制御（Ethical Precision）」の実装指針を提示するものである。
従来のPZT方式に対し，高電流・高温・Pb含有という制約を排除しつつ，
生体分子を損なわない\emph{穏やかな駆動}を実現した。

提案構成は，Si(100)基板上に
SiN$_x$(0.8\,µm)/ALD-Al$_2$O$_3$(60\,nm)/Pt-Ti(100/20\,nm)
を積層した静電アクチュエータであり，
$\leq$400\,\si{\celsius} の低温プロセスにより CMOS BEOL との整合性を確保した。
3.3\,Vロジック＋45\,V高電圧ドライバ構成下で，
膜変位 $\sim$0.1\,µm，圧力 $\sim$50\,kPa にて
1.3\,pL滴を 2–5\,m/s の速度で低衝撃吐出できることを示した。
さらに，Parylene-HT および PEG-SAM 表面処理により
蛋白吸着を抑制し，DNAおよびBSA活性保持率が $\ge$90\% であることを確認した。
FEM解析と実測結果の整合から，
\textbf{約800\,dpi（31.75\,µmピッチ）}が
電場・膜変位・寄生容量・熱・流体の多領域設計から自然に導かれる
合理設計点であることを明らかにした。
本成果は，Pbフリー・非熱・非接触・生体適合を同時に満たす
新たな精密駆動技術の枠組みを提示するものである。

\medskip
\noindent\textbf{Abstract}:~
This paper presents a \emph{lead-free, low-temperature electrostatic thin-film MEMS actuator}
for bio-inkjet applications, establishing a design methodology for \emph{Ethical Precision}—
precision that minimizes physical and biological stress.
Unlike PZT-based systems constrained by high current, high-temperature sintering, and lead toxicity,
the proposed electrostatic approach enables gentle actuation suitable for biomolecule handling.

The actuator consists of a Si(100) substrate with
a SiN$_x$(0.8\,µm)/ALD-Al$_2$O$_3$(60\,nm)/Pt-Ti(100/20\,nm) stack,
fabricated through a $\leq$400\,\si{\celsius} CMOS-compatible process.
Under a 3.3\,V logic + 45\,V high-voltage driver configuration,
a diaphragm displacement of $\sim$0.1\,µm and pressure of $\sim$50\,kPa
enable \textbf{$\sim$1.3\,pL} droplets to be ejected at 2–5\,m/s with minimal impulse.
Surface coatings of Parylene-HT and PEG-SAM suppress protein adsorption,
maintaining DNA/BSA activity above 90\%.
Finite element analysis and experimental validation reveal
an optimal \textbf{$\sim$800\,dpi (31.75\,µm pitch)} design point,
emerging naturally from the co-optimization of electric field,
mechanical compliance, capacitance, and fluid dynamics.
The results demonstrate a Pb-free, non-thermal, non-contact, and bio-compatible
MEMS actuation platform representing a shift toward sustainable precision engineering.
\end{abstract}

\begin{IEEEkeywords}
Electrostatic MEMS actuator, Bio-inkjet, Si(100), ALD-Al$_2$O$_3$, SiN$_x$, Parylene-HT, PEG-SAM, Lead-free, CMOS-compatible, Ethical Precision
\end{IEEEkeywords}
