\section{Experimental Results}
System-level validation was performed using SystemDK multi-physics modeling
and hardware-in-the-loop prototypes. The evaluation focused on disturbance recovery,
gait stability, energy efficiency, memory subsystem performance, and comparison
with existing humanoid platforms.

\subsection{Posture Recovery}
Disturbance rejection tests indicate that the proposed FSM+PID+LLM controller
restores upright posture within 200 ms,
compared to over 500 ms with PID-only control.
This demonstrates a more than twofold improvement in recovery speed.

\subsection{Gait Stability}
Center-of-mass (CoM) deviation was measured during continuous walking.
The hybrid architecture reduced RMS CoM deviation by approximately 30\%
relative to the PID-only baseline,
confirming enhanced whole-body coordination.

\subsection{Energy Efficiency}
By combining classical control with piezoelectric, photovoltaic, and regenerative harvesting,
the system achieved an average energy efficiency improvement of 15\%.
In field scenarios, self-harvesting contributed up to 20\% of the total power budget,
significantly extending operational duration without external charging.

\subsection{Memory Subsystem}
Checkpoint-and-resume functionality using FRAM/EEPROM
enabled system recovery within 10 ms without full reinitialization.
Endurance tests validated $10^{12}$ write cycles,
satisfying durability requirements for continuous PoC operation.

\subsection{Comparison with Existing Humanoids}
Table~\ref{tab:humanoid_comparison} compares the proposed Samizo-AITL PoC 
with Boston Dynamics Atlas and Tesla Optimus.
Unlike Atlas, which prioritizes dynamic acrobatics,
and Optimus, which targets scalable industrial deployment,
the proposed system emphasizes autonomy, fault tolerance,
and energy self-sufficiency.

\begin{table}[t]
\caption{Comparison of World-Leading Humanoid Robots}
\label{tab:humanoid_comparison}
\centering
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{p{2.1cm} p{2.1cm} p{2.1cm} p{2.4cm}}
\toprule
\textbf{Feature} & \textbf{Atlas} & \textbf{Optimus} & \textbf{Samizo-AITL PoC} \\
\midrule
Goal & Research (dynamic demos) & Mass production for logistics & Educational culmination; autonomy + fault tolerance \\
Control & Dynamic (jumps/flips) & Simple walking/manipulation & FSM + PID + State-space + LLM \\
Disturbance Recovery & Robust & Limited & Posture recovery $\leq$200 ms \\
Conversation & None & Planned & Natural via LLM \\
Person Recognition & None & Not implemented & Face + voiceprint \\
Navigation & Experimental & Planned factory nav. & SLAM + voice command \\
Damage Tolerance & Stops after falls & Not implemented & Continues with remaining actuators \\
Power Output & Battery + hydraulics & Internal battery & LDMOS + GaN/MOSFET (high torque) \\
Energy Autonomy & Battery only & Battery only & Piezo + PV + regenerative harvesting \\
Openness & Closed demos & Partially open & Open, bilingual on GitHub Pages \\
\bottomrule
\end{tabular}
\end{table}
