\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

% ===== Packages =====
\usepackage{amsmath,amssymb}
\usepackage{siunitx}
\usepackage{graphicx}
\usepackage{cite}
\usepackage[hidelinks]{hyperref}
\usepackage{tikz}
\usepackage{booktabs}

\sisetup{detect-all=true}

% ===== Title =====
\title{Low-Cost Integration of 1.8-V FeFET on 0.18-\(\mu\)m CMOS:\\
+1 Mask and a Single ALD Tool, with Reliability Assessment}

\author{%
Shinichi Samizo\\
\emph{Independent Semiconductor Researcher}\\
Former Engineer at Seiko Epson Corporation\\
Email: \texttt{shin3t72@gmail.com}\\
GitHub: \url{https://github.com/Samizo-AITL}
}

\begin{document}
\maketitle

% ===== Abstract =====
\begin{abstract}
Ferroelectric FETs (FeFETs) are promising CMOS–compatible embedded NVMs. 
This paper demonstrates a \SI{1.8}{V} FeFET module integrated on a legacy \SI{0.18}{\micro m} CMOS process with only \textbf{one additional mask} and \textbf{a single ALD tool}. 
Fabricated devices show endurance exceeding \(10^{5}\) program/erase cycles and retention longer than 10 years at \SI{85}{\celsius}. 
Reliability was characterized on FeCAP/FeFET structures: time–zero dielectric breakdown (TZDB), time–dependent dielectric breakdown (TDDB), endurance, and retention. 
The approach provides a cost–effective path to extend mature–node lifetimes and to enable embedded NVM for automotive/industrial/IoT, while high–temperature retention remains the key limiter.
\end{abstract}

% ===== Sections =====
\section{Introduction}
FeFETs based on HfO\(_2\) have gained traction as CMOS–compatible NVMs. 
Most prior work targets advanced nodes; however, mature nodes (\(\sim\)\SI{0.18}{\micro m}) remain widely used in automotive/industrial markets where long supply lifetimes and low cost are critical.  
\textbf{This work contributes:}
(i) a \textbf{+1 mask} low–cost module,  
(ii) only \textbf{one ALD tool} added to the line,  
(iii) a yield–friendly \emph{SRAM+FeFET} system usage model, and  
(iv) comprehensive reliability evidence on FeCAP/FeFET.

\section{Process Integration}
Baseline is a \SI{0.18}{\micro m} CMOS platform (\SI{1.8}{V} core, optional \SI{3.3}{V} I/O). 
The FeFET module is inserted after poly definition and salicide/RTA, requiring minimal line modification.

\subsection{Process Flow}
\begin{figure}[t]
\centering
\begin{tikzpicture}[node distance=0.8cm, every node/.style={draw, rectangle, minimum width=2.5cm, minimum height=0.8cm}]
\node (a) {FEOL (Co RTA)};
\node (b) [below=of a] {ALD: Al\(_2\)O\(_3\) (1--2 nm)};
\node (c) [below=of b] {ALD: HZO (8--12 nm)};
\node (d) [below=of c] {TiN (30--50 nm)};
\node (e) [below=of d] {RTA (450--500$^\circ$C)};
\node (f) [below=of e] {BEOL};
\draw[->] (a)--(b);
\draw[->] (b)--(c);
\draw[->] (c)--(d);
\draw[->] (d)--(e);
\draw[->] (e)--(f);
\end{tikzpicture}
\caption{Process flow of FeFET integration.}
\label{fig:proc_flow}
\end{figure}

\subsection{Cross Section}
\begin{figure}[t]
\centering
\begin{tikzpicture}
\draw[fill=gray!30] (0,0) rectangle (3,0.4) node[midway]{TiN 30--50 nm};
\draw[fill=gray!10] (0,0.4) rectangle (3,0.8) node[midway]{HZO 8--12 nm};
\draw[fill=gray!20] (0,0.8) rectangle (3,1.0) node[midway]{Al\(_2\)O\(_3\) 1--2 nm};
\draw[fill=white] (0,1.0) rectangle (3,1.4) node[midway]{p-Si substrate};
\end{tikzpicture}
\caption{Cross section of HZO/Al\(_2\)O\(_3\)/TiN stack.}
\label{fig:cross_section}
\end{figure}

\section{Devices and Methods}
Test structures include FeCAPs (flat/comb) and \SI{100}{\micro m}\(\times\)\SI{100}{\micro m} FeFET cells. 
Programming used \(\pm\)2.3–\SI{2.7}{V}, \SI{1}{–}\SI{50}{\micro s} pulses. 
Keysight B1500A and a manual prober were used.  

Protocols:  
TZDB: DC ramp \(\approx \SI{0.1}{V/s}\) at RT–\SI{125}{\celsius}.  
TDDB: constant–voltage stress at 2.3/2.5/2.7 V, \SI{85}{\celsius} and \SI{125}{\celsius}; Weibull fitting.  
Endurance: \(\pm\SI{2.5}{V}\), \SI{10}{\micro s}, \SI{10}{kHz} up to \(10^{5}\) cycles.  
Retention: \SI{25}{\celsius}, \SI{85}{\celsius}, \SI{125}{\celsius}, with Arrhenius extrapolation.

\section{Results}
\subsection{TZDB}
Breakdown statistics indicate early–failure tails due to defects (Fig.~\ref{fig:tzdb}).

\subsection{TDDB (Weibull/Arrhenius)}
Weibull plots yield \(\beta \approx 1.3\).  
Arrhenius analysis gives activation energies consistent with oxygen–vacancy diffusion:
\(E_a \approx \SI{0.78}{eV} @ 2.3~\mathrm{V}\),
\(\SI{0.84}{eV} @ 2.5~\mathrm{V}\),
\(\SI{0.88}{eV} @ 2.7~\mathrm{V}\).

\begin{table}[t]
\centering
\caption{Extracted TDDB scales \(\eta\).}
\begin{tabular}{@{}ccc@{}}
\toprule
Stress & Temp & \(\eta\) [s] \\
\midrule
2.3 V & \SI{85}{\celsius}  & \(2.7\times 10^{3}\) \\
2.3 V & \SI{125}{\celsius} & \(5.1\times 10^{4}\) \\
2.5 V & \SI{85}{\celsius}  & \(1.5\times 10^{3}\) \\
2.5 V & \SI{125}{\celsius} & \(2.8\times 10^{4}\) \\
2.7 V & \SI{85}{\celsius}  & \(8.2\times 10^{2}\) \\
2.7 V & \SI{125}{\celsius} & \(1.5\times 10^{4}\) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Endurance}
Up to \(10^{5}\) cycles verified; window shrinks 20–30\%.  
\(\Delta V_\mathrm{th}(N)=1.12-0.05\log_{10}N\).

\subsection{Retention}
Arrhenius extrapolation with \(E_a \approx \SI{1.1}{eV}\) predicts  
\(\gt 10^2\) years @ \SI{25}{\celsius},  
\(\gt 10\) years @ \SI{85}{\celsius},  
months @ \SI{150}{\celsius}.

\section{System Architecture (SRAM + FeFET)}
\begin{figure}[t]
\centering
\begin{tikzpicture}[node distance=1.2cm, every node/.style={draw, rectangle, minimum width=2.2cm, minimum height=0.8cm}]
\node (logic) {Logic (1.8V)};
\node (sram) [below=of logic] {SRAM};
\node (fefet) [below=of sram] {FeFET NVM};
\node (pump) [right=of fefet, xshift=3cm] {Charge Pump $\pm$2.5V};
\node (io) [above=of pump] {I/O \& AMS (3.3V)};
\draw[->] (logic)--(sram);
\draw[->] (sram)--(fefet);
\draw[->] (fefet)--(pump);
\draw[->] (pump)--(io);
\end{tikzpicture}
\caption{System architecture with SRAM + FeFET backup.}
\label{fig:system}
\end{figure}

\begin{figure}[t]
\centering
\begin{tikzpicture}[node distance=2.2cm, every node/.style={draw, rectangle, minimum width=2.2cm, minimum height=0.8cm}]
\node (sram) {SRAM};
\node (ctrl) [right=of sram] {Backup Ctrl};
\node (fefet) [right=of ctrl] {FeFET NVM};
\draw[->] (sram)--(ctrl) node[midway,above]{Backup};
\draw[->] (ctrl)--(fefet);
\draw[->] (fefet)--(ctrl) node[midway,below]{Restore};
\draw[->] (ctrl)--(sram);
\end{tikzpicture}
\caption{Backup/restore flow between SRAM and FeFET.}
\label{fig:backup}
\end{figure}

\section{Discussion}
The HZO/Al\(_2\)O\(_3\)/TiN stack provides sufficient reliability for industrial/consumer NVM. 
For high–temperature automotive, improvements are required: IL optimization, crystallinity control, refresh/rewrite, ECC.

\section{Conclusion}
We realized an FeFET module on \SI{0.18}{\micro m} CMOS with one extra mask and one ALD tool. 
Devices exhibit \(>\!10^{5}\) cycles and \(>\!10\) years retention at \SI{85}{\celsius}. 
The method extends mature–node lifetime and enables cost–effective embedded NVM for automotive/industrial/IoT.

\section*{Acknowledgment}
The author thanks collaborators for helpful discussions.

% ===== References =====
\begin{thebibliography}{99}
\bibitem{Boscke2011} T.~Böscke \emph{et al.}, Appl. Phys. Lett., vol.~99, p.~102903, 2011.
\bibitem{Mueller2012} J.~Müller \emph{et al.}, Appl. Phys. Lett., vol.~99, p.~112901, 2012.
\bibitem{Mikolajick2019} T.~Mikolajick \emph{et al.}, J. Appl. Phys., vol.~125, p.~204103, 2019.
\bibitem{Mueller2015} J.~Müller \emph{et al.}, IEEE TED, vol.~62, no.~12, pp.~4158--4166, 2015.
\bibitem{Park2020} J.~Park \emph{et al.}, IEEE EDL, vol.~41, no.~5, pp.~711--714, 2020.
\bibitem{Nakamura2003} H.~Nakamura \emph{et al.}, IEEE TDMR, vol.~3, no.~4, pp.~132--136, 2003.
\bibitem{Yamazaki2018} K.~Yamazaki \emph{et al.}, Jpn. J. Appl. Phys., vol.~57, 04FB07, 2018.
\end{thebibliography}

% ===== Biography =====
\section*{Biography}
\noindent
\textbf{Shinichi Samizo} has over 25 years of experience in semiconductor process integration and actuator development. 
After studying control theory and EM modeling, he joined Seiko Epson in 1997 and worked on \SI{0.35}{\micro m}--\SI{0.18}{\micro m} CMOS logic/memory/HV integration, DRAM, and LCD drivers. 
Later he contributed to PZT actuator development and the PrecisionCore inkjet head. 
He is currently an independent researcher, publishing educational materials via the ``Project Design Hub''.
\end{document}
