\documentclass[conference]{IEEEtran}

% ---------- Robust UTF-8 & fonts (pdflatex / xelatex 両対応) ----------
\usepackage{iftex}
\ifPDFTeX
  \usepackage[utf8]{inputenc}      % ← pdflatex で UTF-8 を確実に
  \usepackage[T1]{fontenc}
  % IEEEtran と相性の良い Times 系
  \usepackage{newtxtext,newtxmath}
\else
  \usepackage{fontspec}            % xelatex / lualatex
  % Times 互換フォント（TU/ptm 警告を回避）
  \setmainfont{TeX Gyre Termes}
  \setsansfont{TeX Gyre Heros}
  \setmonofont{TeX Gyre Cursor}
\fi

% ---------- Common packages ----------
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{cite}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage{booktabs}

% ---------- TikZ / pgfplots ----------
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta}  % right=.. of .. を使う
\tikzset{
  blk/.style={
    rectangle,draw,rounded corners,
    minimum height=9mm,minimum width=26mm,
    align=center,fill=blue!8
  }
}

\usepackage{pgfplots}
\pgfplotsset{compat=1.18}

% ====================================================================
\begin{document}

\title{Pb-free ScAlN MEMS Array Integrated with 65 nm SiGe CMOS via System-in-Package for Medical Ultrasonic Sensors}

\author{
  \IEEEauthorblockN{Shinichi Samizo}
  \IEEEauthorblockA{Independent Semiconductor Researcher\\
  Former Engineer at Seiko Epson Corporation\\
  Email: \href{mailto:shin3t72@gmail.com}{shin3t72@gmail.com}\\
  GitHub: \url{https://github.com/Samizo-AITL}}
}

\maketitle

\begin{abstract}
Conventional medical ultrasonic devices have been dominated by PZT (Pb(Zr,Ti)O$_3$)~\cite{akata2009pzt}. However, Pb-containing materials face strict regulatory restrictions (EU RoHS, REACH, FDA) in in-body medical applications. This work proposes a Pb-free alternative based on ScAlN MEMS arrays, integrated with 65~nm SiGe CMOS using System-in-Package (SiP) technology.
The ScAlN MEMS array is designed for 10--50~MHz operation with $\lambda/2$ pitch for high-resolution imaging~\cite{akrout2018scaln}. The SiGe CMOS front-end integrates LNA, VGA, and ADC, enabling detection of microvolt-level signals. The SiP approach ensures yield separation, short interconnects, and hermetic sealing for medical reliability. Finite element and circuit simulations indicate adequate sensitivity and beam directivity at 20--40~MHz, LNA noise figure $<2$~dB, and compact, reliable packaging via flip-chip SiP. Pb-free ScAlN arrays with 65~nm SiGe CMOS via SiP form a practical path for next-generation high-resolution medical ultrasonic sensors.
\end{abstract}

\begin{IEEEkeywords}
ScAlN MEMS, Pb-free piezoelectrics, System-in-Package (SiP), SiGe CMOS, Medical ultrasound, Ultrasonic imaging
\end{IEEEkeywords}

% ====================================================================
\section{Introduction}
Medical ultrasonic imaging is widely used in ophthalmology, vascular diagnosis, dermatology, and implantable monitoring. Traditional devices have been dominated by PZT due to superior piezoelectric properties~\cite{akata2009pzt}, but Pb toxicity limits in-body use under EU RoHS, REACH, and FDA regulations. Scandium-doped AlN (ScAlN) has emerged as a Pb-free candidate with CMOS compatibility, high-$Q$, and industrial adoption in RF BAW/XBAR filters~\cite{akrout2018scaln}. We propose ScAlN MEMS arrays co-integrated with 65~nm SiGe CMOS via SiP for medical ultrasound.

% ====================================================================
\section{Background and Contributions}
\subsection{Background}
PZT transducers provide excellent $d_{33}$ and coupling but contain Pb and are not preferred for in-body devices~\cite{akata2009pzt}. CMUT/PMUT enable MEMS-CMOS co-integration, yet often require high bias and face fluid-reliability challenges~\cite{khuri2009cmut}. ScAlN is a promising Pb-free piezoelectric with proven manufacturability and CMOS process compatibility~\cite{akrout2018scaln}. Combining ScAlN arrays with low-noise SiGe readout via SiP offers a realistic, compliant route to high-resolution ultrasound.

\subsection{Contributions}
\begin{itemize}
  \item Pb-free ScAlN MEMS array architecture for 10--50~MHz imaging.
  \item Heterogeneous integration with 65~nm SiGe CMOS via SiP to minimize interconnect and improve SNR.
  \item FEM/circuit examples indicating adequate sensitivity, beam directivity, and low-noise detection at 20--40~MHz.
  \item Application considerations in ophthalmology~\cite{pavlin2009ubm}, IVUS~\cite{foster2000ivus}, dermatology, and implantables.
\end{itemize}

% ====================================================================
\section{System Concept}
\subsection{ScAlN MEMS Array}
Operation: 10--50~MHz. Channels: 64--256 with $\lambda/2$ pitch. Structures: PMUT-like stacks or BAW/XBAR cavities.

\subsection{SiGe CMOS Front-end}
65~nm SiGe BiCMOS with low NF ($<2$~dB). Integrated LNA, VGA, ADC, and T/R switch to sense $\mu$V-level signals.

\subsection{System-in-Package Integration}
Flip-chip bonding integrates MEMS and CMOS dice. Benefits: yield separation, short interconnects, and hermetic sealing.

\begin{figure}[t]
\centering
\begin{tikzpicture}[node distance=10mm, auto, thick]
\node[blk]              (mems) {ScAlN MEMS Array\\(10--50 MHz)};
\node[blk, below=10mm of mems] (lna)  {SiGe LNA/VGA\\(65 nm)};
\node[blk, below=10mm of lna]  (adc)  {ADC\\(12--14 bit, 50--100 MS/s)};
\node[blk, below=10mm of adc]  (bf)   {Beamformer/\\Digital Core};

\draw[->] (mems)--(lna);
\draw[->] (lna)--(adc);
\draw[->] (adc)--(bf);
\end{tikzpicture}
\caption{System architecture (ScAlN MEMS + 65~nm SiGe via SiP).}
\label{fig:arch}
\end{figure}

% ====================================================================
\section{FEM Analysis Methodology}

Finite element simulations were performed to evaluate the
ScAlN MEMS array performance. The device stack (Sc$_{0.2}$Al$_{0.8}$N
piezoelectric layer, electrodes, and supporting membrane) was modeled
with piezoelectric solid mechanics, while the surrounding water domain
was represented by acoustic pressure elements with perfectly matched
layers (PML) to emulate open boundaries.

Boundary conditions were:
\begin{itemize}
  \item \textbf{Transmit:} Top electrode driven by 1~V AC, bottom electrode grounded.
  \item \textbf{Receive:} Electrodes left open, and a 1~Pa plane acoustic
        wave was applied at the fluid boundary.
  \item \textbf{Support:} Membrane rim mechanically fixed.
\end{itemize}

Material constants were taken from literature for
ScAlN ($\rho\!\approx\!3200$~kg/m$^3$, $\varepsilon_r\!\approx\!11$,
$e_{33}\!\approx\!1.3$~C/m$^2$). Frequency-domain sweeps from
10--50~MHz were conducted to extract impedance, displacement fields, and
acoustic pressure radiation. Effective electromechanical coupling
$k_\mathrm{eff}^2$ was calculated from resonance and anti-resonance
frequencies. The open-circuit receive voltage was combined with the SiGe
LNA noise model ($<2$~nV/$\sqrt{\mathrm{Hz}}$) to estimate overall
system SNR.

% ====================================================================
\section{Simulation Results}

FEM analysis revealed resonances near 20, 30, and 40~MHz corresponding
to the designed membrane/cavity modes. The simulated electrical
impedance exhibited clear series and parallel resonances; the extracted
effective coupling was $k^2_{\mathrm{eff}}=2$--4\% based on the
frequency split. In the transmit case (1~V$_\mathrm{AC}$ drive), the
predicted on-axis pressure reached the order of tens of kPa at 1~mm
depth in water, sufficient for ophthalmic and IVUS ranges. In the
receive case (1~Pa plane wave), the open-circuit voltage was in the
tens-of-$\mu$V range. Combined with the 65~nm SiGe front-end
(input-referred noise $<2$~nV/$\sqrt{\mathrm{Hz}}$), the system SNR
exceeded 60~dB across 20--40~MHz (Fig.~\ref{fig:snr}).

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{semilogyaxis}[
  width=\columnwidth, height=5.0cm,
  xlabel={Frequency (MHz)}, ylabel={$|Z|$ (Ohm)},
  xmin=10, xmax=50, ymin=50, ymax=1e5,
  grid=both, tick align=inside,
  legend style={at={(0.02,0.98)},anchor=north west}
]
% サンプル座標（必要に応じて実データに置換）
\addplot+[thick,mark=*] coordinates {
 (10,8000) (12,5000) (14,3000) (16,1200) (18,600)
 (19,300) (20,180) (21,280) (22,700) (24,2000)
 (26,4000) (28,7000) (29,9000) (30,12000) (31,9000)
 (32,6000) (34,3000) (36,1200) (38,600)
 (39,320) (40,200) (41,300) (42,700) (45,3000) (48,7000) (50,10000)
};
\legend{$|Z(f)|$}
\end{semilogyaxis}
\end{tikzpicture}
\caption{Representative impedance magnitude showing series/parallel
resonances near 20, 30, and 40~MHz.}
\label{fig:imp}
\end{figure}

\begin{figure}[t]
\centering
\begin{tikzpicture}[scale=0.9]
% 水 + PML
\fill[blue!6] (-3,1.6) rectangle (3,3.0);     % Water
\fill[blue!12,opacity=0.7] (-3,3.0) rectangle (3,3.6); % PML

% 膜/層
\fill[gray!25] (-2.2,1.2) rectangle (2.2,1.6);      % ScAlN
\fill[gray!45] (-2.2,1.6) rectangle (2.2,1.65);     % Top electrode
\fill[gray!45] (-2.2,1.15) rectangle (2.2,1.2);     % Bottom electrode

% 支持リム (固定)
\fill[gray!50] (-2.4,0.9) rectangle (-2.2,1.8);
\fill[gray!50] ( 2.2,0.9) rectangle ( 2.4,1.8);

% 基板
\fill[brown!20] (-3,0.9) rectangle (3,1.15);

% ラベル
\node at (0,3.3) {PML};
\node at (0,2.3) {Water domain};
\node at (0,1.4) {ScAlN layer};
\node[anchor=east] at (-2.4,1.35) {Fixed rim};
\node[anchor=west] at (2.4,1.35) {Fixed rim};

% 電気境界
\draw[->] (-2.8,1.625) -- (-2.25,1.625);
\node[anchor=east] at (-2.8,1.625) {Top electrode: 1 V (Tx) / Open (Rx)};
\draw[->] (-2.8,1.18) -- (-2.25,1.18);
\node[anchor=east] at (-2.8,1.18) {Bottom electrode: GND};

% 音響境界
\draw[->] (2.8,2.7) -- (2.3,2.7);
\node[anchor=west] at (2.8,2.7) {Plane wave 1 Pa (Rx)};
\draw[dashed] (-3,3.0) -- (3,3.0);

\end{tikzpicture}
\caption{FEM model sketch: stack, boundary conditions, and acoustic PML.}
\label{fig:mode}
\end{figure}

% ====================================================================
\section{Application Scenarios}
\begin{itemize}
  \item \textbf{Ophthalmology:} 20--40~MHz anterior eye imaging~\cite{pavlin2009ubm}.
  \item \textbf{Vascular IVUS:} 30--40~MHz catheter arrays~\cite{foster2000ivus}.
  \item \textbf{Dermatology:} 10--20~MHz skin/tumor imaging.
  \item \textbf{Implantables:} Miniaturized SiP with telemetry.
\end{itemize}

% ====================================================================
\section{Discussion}
Versus PZT, ScAlN offers CMOS compatibility and Pb-free compliance at lower $d_{33}$. Versus CMUT/PMUT~\cite{khuri2009cmut}, ScAlN needs lower bias and shows better fluid reliability. SiP yields separation, short interconnects, and hermeticity versus monolithic integration.

% ====================================================================
\section{Conclusion}
Pb-free ScAlN MEMS arrays integrated with 65~nm SiGe CMOS via SiP provide: (i) regulatory compliance, (ii) adequate 20--50~MHz resolution, (iii) low-noise detection via SiGe, and (iv) scalable, reliable packaging. This is a practical, competitive path for next-generation medical ultrasonic sensors.

% ====================================================================
\section*{Acknowledgment}
The author thanks colleagues and collaborators in semiconductor device research and MEMS development.

\bibliographystyle{IEEEtran}
\bibliography{references}

% ====================================================================
\section*{Author Biography}
\textbf{Shinichi Samizo} received the M.S. degree in Electrical and Electronic Engineering from Shinshu University, Japan. He joined Seiko Epson Corporation in 1997, engaging in semiconductor device process development including 0.25--0.18~$\mu$m CMOS, HV-CMOS, DRAM, FeRAM. He also contributed to inkjet MEMS process development and thin-film piezo actuator design, leading to the productization of PrecisionCore printheads. His expertise covers semiconductor devices (logic, memory [DRAM/FeRAM/SRAM], high-voltage mixed integration), inkjet actuators, and AI-based control education.

\end{document}
