\documentclass[conference]{IEEEtran}

% ===== Packages =====
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{cite}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,fit,calc,shapes.geometric,decorations.pathreplacing}
\usepackage[hidelinks]{hyperref}
\urlstyle{same}

% ===== Title & Author =====
\title{Lead-free Bio-Inkjet Printing with Bulk KNN Actuators and AITL-based Adaptive Control}

\author{
  \IEEEauthorblockN{Shinichi Samizo}
  \IEEEauthorblockA{Independent Semiconductor Researcher\\
  Former Engineer at Seiko Epson Corporation\\
  Email: \href{mailto:shin3t72@gmail.com}{shin3t72@gmail.com}\\
  GitHub: \url{https://github.com/Samizo-AITL}}
}

\begin{document}
\maketitle

% ===== Abstract =====
\begin{abstract}
This paper proposes a biocompatible inkjet printing architecture based on
lead-free piezoelectric actuators, specifically bulk KNN (K,Na)NbO$_3$, combined
with chip-on-film (COF) driver ICs, silicon cavity integration, and
an AITL (Adaptive Intelligent Three-Layer) supervisory control framework.
Unlike conventional PZT-driven industrial heads, the KNN approach supports
biocompatibility and Pb-free compliance, while moderate performance
(3--5 pL droplets, $10^6$ shots, $\pm50$ V operation) satisfies bioprinting needs.
Beyond actuation, the electromechanical response of KNN is leveraged for
\emph{self-diagnosis}: detecting missing droplets and estimating ink viscosity.
This diagnostic data feeds into an AITL architecture---an inner PID loop for
real-time stability, an FSM layer for mode transitions and compensation,
and an LLM-driven supervisory layer for adaptive redesign of control rules.
We present the architecture, process flow, and representative biomedical
applications, highlighting the potential of KNN as both actuator and sensor.
\end{abstract}

% ===== Sections =====
\section{Introduction}
\input{sections/intro}

\section{Background: Pb-free Piezoelectrics}
\input{sections/background}

\section{Proposed Bio-IJ Architecture}
\input{sections/proposal}

\section{Self-Diagnosis and Feedback Control}
The bulk KNN actuator can also function as a proxy sensor by monitoring
its electromechanical response during drive pulses.
\subsection{Dot-Missing Detection and Compensation}
By observing charge/displacement signatures, missing droplets are detected
without external optics. An FSM controller triggers neighboring-nozzle
compensation, preserving array-level print reliability.
\subsection{Viscosity Estimation and Drive Adaptation}
Viscosity changes appear as altered dynamic response.
The PID loop adjusts drive voltage/pulsewidth in real time.
This closes the loop between fluid rheology and actuator operation.

\section{AITL Control Framework}
The Adaptive Intelligent Three-Layer (AITL) framework integrates:
\begin{itemize}
\item \textbf{Inner PID}: ensures real-time stability of droplet ejection.
\item \textbf{Intermediate FSM}: supervises nozzle states, detects anomalies,
and switches modes (normal, compensation, cleaning).
\item \textbf{Outer LLM}: analyzes long-term trends, re-identifies PID gains,
and redefines FSM transitions when performance degrades.
\end{itemize}
A schematic is shown in Fig.~\ref{fig:aitl}.
\begin{figure}[t]
\centering
\begin{tikzpicture}[
  block/.style={draw, rounded corners, minimum width=2.4cm, minimum height=8mm, align=center},
  line/.style={-{Latex[length=2.2mm,width=1.4mm]}}
]
\node[block] (pid) {PID Loop\\(Droplet Stability)};
\node[block, above=of pid] (fsm) {FSM Layer\\(Nozzle State Control)};
\node[block, above=of fsm] (llm) {LLM Layer\\(Adaptive Redesign)};
\draw[line] (fsm) -- (pid);
\draw[line] (llm) -- (fsm);
\end{tikzpicture}
\caption{AITL three-layer architecture for Bio-IJ adaptive control.}
\label{fig:aitl}
\end{figure}

\section{Applications}
\input{sections/applications}

\section{Conclusion}
\input{sections/conclusion}

% ===== References =====
\nocite{Saito2004KNN,Dubois2019BioIJ}
\bibliographystyle{IEEEtran}
\bibliography{refs}

% ===== Biography =====
\section*{Author Biography}
\textbf{Shinichi Samizo} received the M.S. degree in Electrical and Electronic
Engineering from Shinshu University, Japan. He worked at Seiko Epson
Corporation as an engineer in semiconductor memory and mixed-signal
device development, and contributed to inkjet MEMS actuators and
PrecisionCore printhead technology. He is currently an independent
semiconductor researcher focusing on process/device education, memory
architecture, and AI system integration. Contact:
\href{mailto:shin3t72@gmail.com}{shin3t72@gmail.com}.
\end{document}
